---
title: "New_DYNATE_Tutorial"
author: "Jintong Zhao"
date: "2023-05-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
all_times <- list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res <- difftime(Sys.time(), now, units = "secs")
      all_times[[options$label]] <<- res
    }
  }
}))
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
load("/Users/mac/Desktop/RA/snp_dat.rda")
```

## Setup the DYNATE Object 

For this tutorial, we will use `DYNATE` package to analyze a subset of ALS data contains 100000 rows (10138 samples, 7504 controls and 2634 cases) and total number of rare variants that all patients have are 66046.

```{r input data}
library(dplyr)
library(DYNATE)
library(data.table)
library(reshape2)
library(tibble)
# Load the ALS dataset
load("/Users/mac/Desktop/RA/ALS_Preprocessed_Data_Euro_Sample.RData")
raw = ALS_data_SNV_no_LoF_euro
```

We start by reading in the data. We only select the domain that have at least 5 SNPs records residing in them. i.e. a sample with 2 SNPs will be considered as 2 records. We first choose the domains that satisfy the criteria and then choose a subgroup of domains as the input data.

```{r}
set.seed(111) 
raw <- raw%>%dplyr::rename(Sample.Name=ID)
#filter begin
raw <- raw%>%group_by(subRVIS.Domain.Name)%>%
  mutate(len=n())%>%
  ungroup()%>%
  filter(len>=5)%>%
  arrange(loc_adj)%>%
  mutate(snpID0=Variant.ID,snpID=rleid(loc_adj),Sample.Type=as.character(Sample.Type))%>%
  arrange(-desc(Sample.Type),Sample.Name)%>%
  mutate(Sample.ID=rleid(Sample.Name)) %>% arrange(subRVIS.Domain.Name)%>%
  mutate(domainID=rleid(subRVIS.Domain.Name))
  
#filter end 
snp_dat1 = raw %>% select("Sample.Type","Sample.ID","snpID","domainID")
#snp_dat1 = as.data.frame(lapply(snp_dat1, sort))
snp_dat2 = snp_dat1[1:100000, ]
snp_dat2<- rename(snp_dat2, Sample.Name = Sample.ID)
snp_dat2$snpID = as.numeric(snp_dat2$snpID)

```

```{r}
sample_info = snp_dat2 %>% select("Sample.Name","Sample.Type")%>% unique()
```
The dimension of the `sample_info` data is $10138 \times 2$.

This data has 10138 rows and 2 columns with name `Sample.Name`, `Sample.Type`. Each row respresents  an individual patient. 

Here is the first 10 rows of the `sample_info` data.

```{r display data}
head(sample_info,10)
```

`Sample.Name`: A sample unique identifier for each patient. The data contains 100000 samples selected from the raw ALS dataset. Each value indicates the source of the sample data for the row.

`Sample.Type`: Char categorical variable. It indicates the sample in this row is from which treatment group. In this case, there are only two groups: case group or the control group.

Case group (denoted as "case") and control group (denoted as "ctrl") contain 2634 and 7504 samples respectively. The case group is composed of people who have the outcome of interest, while the control group is composed of people who do not have the outcome of interest. 

```{r}
snp_info = snp_dat2 %>% select("Sample.Name","snpID","domainID")
```

The dimension of the `snp_info` data is $100000 \times 3$.

This data has 100000 rows and 3 columns with name `Sample.Name`, `snpID`,`domainID`. Each row respresents a sample specific rare variant (SNP). 

Here is the first 10 rows of the `snp_info` data.

```{r}
head(snp_info,10)
```

`snpID`: A string to respresent SNP. An snp ID is a unique identifier that is assigned to a specific SNP(Single Nucleotide Polymorphism). These IDs are used to refer to specific SNPs in research.

`domainID`: Discrete numeric variable. The data contains 949 domain IDs. It indicates the homology-based protein domain where this row's snp is located in.

From the data above we can see that there are 5 people with snp "117". In total there are 66046 SNPs. But we have 100000 rows, which means that some variants(SNPs) do not appear in only one sample.

A histogram below plots the number of samples who have the rare variant in each snp ID.
For example, when the x axis Snp ID is 263301
, the y axis is 6, which means that the mutation at Snp ID = 263301
 appeared in 6 samples.
```{r snpID, echo=FALSE}
library(ggplot2)
library(gridExtra)

p1 = ggplot(data.frame(y = snp_info$snpID), aes(x = y)) +
  geom_histogram(binwidth = 1, color = "black") +
  labs(title = "The Number of Samples Who Have the Rare Variant in Each Snp ID",x = "Snp ID",y = "Number of Samples") + 
  theme_bw()
p1
```

A histogram below plots the number of SNPs occurring in each domain in all samples
For example, when the x axis Domain ID is 1, the y axis is 10, which means in all samples, a total of 10 variations occurred in domain 1.
```{r domainID, echo=FALSE}
p3 = ggplot(data.frame(y = snp_info$domainID), aes(x = y)) +
  geom_histogram(binwidth = 0.2, color = "black") +
  labs(title = "The Number of Domain Each Sample's Rare Variant From", x = "Domain ID", y = "Value") + 
  theme_bw()
p3
```

```{r include=FALSE}
Test_Leaf1 <- function(snp_info=NULL,sample_info = NULL, thresh_val=10,covars=NULL,teststat="FET"){
  sample_info <- sample_info %>% select("Sample.Name","Sample.Type") %>% unique()
  snp_info <- snp_info %>% select("Sample.Name","snpID","domainID")
  
  # Prove we can gain former input data by using new input data
  snp_dat <- left_join(snp_info, sample_info, by = "Sample.Name")
  struct_map <- construct_leafs(snp_dat=snp_dat,thresh_val=thresh_val)
  total_leaves <- uniqueN(struct_map$L1)
  #Global objects
  D_approx_prev = FD_approx_prev = 0
  
  glm_input <- snp_dat%>%
    dplyr::select(c("Sample.Name","Sample.Type",covars))%>%
    distinct_all()%>%
    arrange(Sample.Name)%>%
    mutate(types=ifelse(Sample.Type=="case",1,0))%>%
    dplyr::select(c("types",covars))
  
  N <- uniqueN(snp_dat$Sample.Name)
  nsnp <- max(as.integer(snp_dat$snpID))
  # Get Gmat case and control
  mat_all <- new("dgTMatrix",
                 i = as.integer(snp_dat$Sample.Name-1),
                 j = as.integer(snp_dat$snpID-1), x=rep(1,nrow(snp_dat)),
                 Dim = c(N, nsnp))
  N1=sum(glm_input$types)
  N0=N-N1
  
  leaf_mat_all <- create_leaf_attribute(mat_all,struct_map)
  leaf_mat_all@x <- ifelse(leaf_mat_all@x>0,1,0)
  
  if(teststat=="FET"){
    #Get marginals - qualifying variants for cases and all
    case_colSums = Matrix::colSums(leaf_mat_all[1:N1,])
    all_colSums = Matrix::colSums(leaf_mat_all)
    
    pvals.1 <- calcFETpval_per_leaf(N1=N1,N0=N0,
                                    case_colSums = case_colSums,
                                    all_colSums = all_colSums,
                                    midp=TRUE)
    pvals.1 <- pmin(pvals.1,1)
    
  } else{
    if(ncol(glm_input)>1){
      #With covariates
      score.test <- ScoreTest_fastSPA_sparse(genomat=leaf_mat_all, #leaf_mat_all is an Nxm matrix of leaf attributes
                                             pheno=glm_input$types,cov=glm_input[,covars],
                                             minmac=1,Cutoff=2)
    } else{
      #Without covariates
      score.test <- ScoreTest_fastSPA_sparse(genomat=leaf_mat_all, #leaf_mat_all is an Nxm matrix of leaf attributes
                                             pheno=glm_input,cov=NULL,
                                             minmac=1,Cutoff=2)
    }
    score.1 <- score.test$Tstat.sign
    pvals.1 <- score.test$p.value
  }
  
  Z.1 <- qnorm(1-pvals.1)
  pvals.1[is.na(pvals.1)] <- 1
  Z.1[is.na(pvals.1)] <- -100
  Z.1 <- pmin(Z.1,100)
  Z.1 <- pmax(Z.1,-100)
  S <- NULL
  #Do not test the leaf with pvals.1=1###########check!!!!
  struct_map0 <- data.table("L1"=seq_along(pvals.1),"pvals"=pvals.1)%>%
    left_join(struct_map,by="L1")%>%filter(pvals!=1)%>%#mutate(L0=L1)%>%
    #mutate(L0=rleid(L1))%>%
    mutate("Test"=teststat)
  
  return(struct_map0)
}


#' DYNATE
#' Function to conduct hierarchical mutiple testing based on the leaf p-values
#' @import tidyverse dplyr
#' @importFrom tibble rowid_to_column
#' @importFrom methods as
#' @importFrom data.table ":=" uniqueN data.table setkey setDT rleid
#' @importFrom stats uniroot pchisq splinefunH na.omit integrate pchisq pnorm qnorm dhyper glm model.matrix
#' @importFrom reshape2 colsplit
#' @import Matrix
#' @param struct_map a data frame with both leaf information and P-value information.
#' @param L maximum number of layers
#' @param alpha desired FDR
#'
#' @return a data frame with testing results.
#' @export
#'
#' @references
#' Li, Xuechan, Anthony Sung, and Jichun Xie. "Distance Assisted Recursive Testing." arXiv preprint arXiv:2103.11085 (2021).
#' Pura, John, et al. "TEAM: A Multiple Testing Algorithm on the Aggregation Tree for Flow Cytometry Analysis." arXiv preprint arXiv:1906.07757 (2019).
#'
#' @examples
#' data("p_leaf")
#'
#' # Set tuning parameters
#' L <- 3 # layer number
#' alpha <- 0.05 # desired FDR
#'
#' # conduct dynamic and hierarchical testing based on the leaf level p values.
#' out <- DYNATE(struct_map=p_leaf,L=L,alpha=alpha)
#' summary(out)
DYNATE6 <- function(struct_map,
                   L=5,
                   alpha=0.05){
  
  total_leaves <- uniqueN(struct_map$L1)
  
  struct_map <- data.table(struct_map)%>%mutate(L0=L1,L1=rleid(L1))
  pvals.1 <- struct_map[,c("pvals","L1")]%>%distinct_all()
  pvals.1 <- c(struct_map$pvals)
  
  # get the estimated mixed node structure
  total_leaves <- uniqueN(struct_map$L1)
  hatn1 <- ceiling(sqrt(total_leaves))
  pn1 <- pvals.1[rank(pvals.1,ties.method="first")==hatn1]
  
  setkey(struct_map,snpID)
  struct_map <- struct_map %>% mutate(hatm1=(pvals<=pn1),pvals1=pvals) %>%
    group_by(L1)%>%mutate(wt=1/n())%>%ungroup()
  
  struct_map_ <- struct_map
  
  Sps <- NULL
  p1s <- p0s <- NULL
  smap_res <- NULL
  S<-NULL
  D_approx_prev = FD_approx_prev = 0
  
  #pvals_list <- list() #add
  
  #Ll <- paste0("L",1) # 添加这一行，定义初始值为 L1
  
  for(l in seq(L)){
    Ll <- paste0("L",l)
    Lm1 <- paste0("L",l-1)
   
    if(l>1){
      #First remove domains with fewer than 2^(l-1) L1 leaves
      removed_map <- struct_map%>%
        group_by(domainID)%>%filter_at(Lm1,any_vars(uniqueN(.)<2))%>%
        ungroup()
      struct_map <- struct_map%>%
        group_by(domainID)%>%filter_at(Lm1,any_vars(uniqueN(.)>=2))%>%
        ungroup()
      
      setDT(struct_map)
      setkey(struct_map, domainID)
      struct_map <- struct_map[, (Ll) := {
        ## modify rle values
        x <- ceiling(rleid(get(Lm1)) / 2)
        n <- uniqueN(get(Lm1))
        if(n %% 2==1){
          x[x == x[.N]] <- x[.N] - 1
        }
        x
      }, by = .(domainID)][, (Ll) := rleid(domainID,get(Ll))] #reassign groups by domain and Ll
      
      struct_map <- struct_map%>%group_by_at(Ll)%>%#mutate(wt=)%>%
        mutate(pvals=stouffer.p2(pvals1,L1,wt))%>%ungroup()
      Sps <- samp.pvals.leaf_L1(struct_map,Ll,p1s=p1s,p0s=p0s)
      
      struct_map <- struct_map %>% mutate(!!paste0("pvals-", l) := pvals) #add
    }
    
    
   # Ll <- paste0("L",l) # 添加这一行，将 Ll 更新为当前循环的 L 值
    
    m.l = uniqueN(data.frame(struct_map)[,Ll])
    
    
    #pvals_cols <- grep("^pvals-", colnames(struct_map)) # add
    #struct_map_pvals <- struct_map[, pvals_cols] # add
  #  pvals.l = data.frame(struct_map_pvals[,l-1]) %>% unique() # add
    #pvals.l = data.frame(struct_map_pvals[,l-1]) # add
    #pvals.l = pvals.l[!duplicated(pvals.l), ] # add
    
    #pvals.l = unlist(pvals.l) # add
    
   pvals.l=struct_map%>%dplyr::select(c("pvals",all_of(Ll)))%>%distinct_all()

  
    
    #Obtain layer specific-threshold
    p.hat = est.p.hat_samp(l=l,
                           D_prev = D_approx_prev,
                           FD_prev = FD_approx_prev,
                           #pvals_l =pvals.l,# add
                           pvals_l=c(pvals.l$pvals),
                           alpha=alpha,
                           alpha1=NULL,
                           Sps=Sps)
    
    
    p.hat.l=p.hat$p.hat
    D_approx_prev = p.hat$D_approx_prev
    FD_approx_prev = p.hat$FD_approx_prev
    if(l==1){
      pvals.1 <- c(pvals.l$pvals)
      S.l1 <- which(pvals.1 <= p.hat$p.hat1)
      if(length(S.l1)==0){
        p1s=min(pvals.1)
        p0s=pvals.1[-which.min(pvals.1)]
      }else{
        p1s=pvals.1[S.l1]
        p0s=pvals.1[-S.l1]
        
        #new (R(1-alpha)), more conservative, 02242022
        p1s1=sort(p1s)[seq(floor(length(p1s)*(1-alpha)))]
        p1s2=sort(p1s)[-seq(floor(length(p1s)*(1-alpha)))]
        p1s=p1s1
        p0s=c(p0s,p1s2)
        #new end
      }
    }
    rej_map <- struct_map%>%filter(pvals<=p.hat.l)%>%mutate(Layer=l)
    smap_res <- smap_res%>%bind_rows(rej_map)
    struct_map <- struct_map%>%filter(pvals>p.hat.l)
    if(nrow(struct_map)==0){next}
  }
  smap_res <- smap_res%>%mutate(L1=L0)%>%
    dplyr::select(-c("pvals","hatm1","wt","L0"))
  
  return(smap_res)
}


############# AUXILIARY FUNCTIONS #################
construct_leafs <- function(snp_dat,
                            thresh_val=10){
  ##Leaf construction
  t1 <- Sys.time()
  #function(thresh, ID, domain_end,snp_end)
  struct_map <- data.table(snp_dat) %>%
    arrange(domainID,snpID)%>%
    mutate_at("Sample.Name",as.character) %>%
    #Coerce ID from factor to character
    group_by(domainID) %>%
    mutate(lastObsFlagDomain = as.integer(row_number() == n())) %>%
    group_by(snpID,.add=TRUE) %>%
    mutate(lastObsFlagSnp = as.integer(row_number()==n())) %>%
    ungroup() %>%
    mutate(num_comb = compute_counts(thresh_val, Sample.Name,
                                     lastObsFlagDomain,lastObsFlagSnp)) %>%
    mutate(num_unique = colsplit(num_comb,",",c(1:2))[,1]) %>%
    mutate(num_all = colsplit(num_comb,",",c(1:2))[,2]) %>%
    mutate(group = cumsum(c(-1L, diff(num_all)) <= 0L)) %>%
    group_by(domainID) %>%
    mutate(group2 = ifelse(group==max(group) & last(num_unique) < thresh_val,
                           max(max(group)-1L,min(group)),group)) %>%
    ungroup() %>%
    mutate(L1 = rleid(group2)) %>%
    dplyr::select(-c(contains("group"),lastObsFlagDomain,num_unique,num_comb,num_all)) %>%
    mutate_at("Sample.Name",as.factor) %>% #coerce ID back to factor
    data.table() #15-37s, based on N0 size
  
  struct_map <- struct_map %>% dplyr::select(c(snpID,L1,domainID)) %>% distinct(snpID,L1,domainID)
  setkey(struct_map,snpID)
  t2 <- Sys.time()
  
  return(struct_map)
}


est.p.hat_samp <- function(l,D_prev,FD_prev,pvals_l,alpha,alpha1=NULL,Sps=NULL){
  
  ##print(paste("l:",l))
  if(is.null(alpha1)) {alpha1=alpha}
  m.l = length(pvals_l)
  
  #Threshold very small p-values based on constant
  p.m = ifelse(m.l==0,1,min(1/(m.l*sqrt(log(m.l))),0.05))
  
  filter <- which(pvals_l<=p.m|pvals_l>alpha1)
  
  if(length(filter)>0){
    p.vec = unique(sort(pvals_l[-filter],decreasing = FALSE))
  }else {
    p.vec = unique(sort(pvals_l,decreasing = FALSE))
  }
  if(length(p.vec)==0){p.vec=p.m}
  
  p.indx = 0
  emp.fdr = 0
  
  addi <- 0
  if(!is.null(Sps)){
    propct <- Sps$sps$ct*Sps$sps$altprop
    delnames <- intersect(c("type","altct","nullct","altprop","ct"),colnames(Sps$sps))
    sps <- Sps$sps%>%dplyr::select(-delnames)
    
    rows=nrow(sps);cols=ncol(sps)
    if(rows==1){
      sps0 <- sapply(sps,function(x){propct%*%outer(x,p.vec,"<=")})
      addi <- mean(sps0)
    }else{
      sps0 <- sapply(sps,function(x){colSums(diag(propct)%*%outer(x,p.vec,"<="))})
      if(length(p.vec)==1){
        addi <- mean(sps0)
      }else{
        addi <- rowMeans(sps0)
      }
      
    }
  }
  
  fdr.num = FD_prev + m.l*p.vec+addi
  fdr.denom = D_prev + sapply(p.vec,function(x){sum(pvals_l<=x)})
  emp.fdr = fdr.num/pmax(fdr.denom,1)
  p.vec <- c(p.m,p.vec)
  index <- max(which(c(0,emp.fdr)<=alpha))
  index1 <- max(which(c(0,emp.fdr)<=alpha1))
  p.hat <- p.vec[index]
  p.hat1 <- p.vec[index1]
  D_approx_prev=fdr.denom[index]
  FD_approx_prev=min(fdr.num[index],D_approx_prev)
  
  return(list("p.hat"=p.hat,"p.hat1"=p.hat1,
              "FD_approx_prev"=FD_approx_prev,
              "D_approx_prev"=D_approx_prev))
}



####### Functions to compute test-statistics

fisher.exact.test <- function(z,midp=TRUE){
  
  x <- z[1]
  sampTot <- z[2]
  pop1Tot <- z[3]
  pop2Tot <- z[4]
  
  lo <- max(0L, sampTot - pop2Tot)
  hi <- min(sampTot, pop1Tot)
  
  support <- lo : hi
  out <- dhyper(support, pop1Tot, pop2Tot, sampTot)
  
  if(midp){
    #mid p-val with minimum likelihood method
    return(sum(out[out < out[x - lo + 1]]) + sum(out[out==out[x-lo+1]])/2)
  } else{
    #minimum likelihood method
    return(sum(out[out <= out[x - lo + 1]]))
  }
}

calcFETpval_per_leaf <- function(N1,N0,case_colSums,all_colSums,
                                 midp=TRUE){
  
  #Order of input is: sample hit, sample size, pop1 size, pop2 size
  cont_tab_summ2x2 <- unname(cbind(case_colSums,all_colSums,N1,N0))
  
  # Apply row-wise
  FET_pvals <- apply(cont_tab_summ2x2, 1,
                     function(z) fisher.exact.test(z,midp = midp))
  
  #Vector of length m.l, where m.l is number of leaves/hypotheses at layer l
  return(FET_pvals)
  
}

compute_counts <- function(thresh, ID, domain_end,snp_end) {
  see_idss <- seen_ids <- NULL
  count <- 0L
  countall <- 0L
  adjust_count <- function(id, domain_end,snp_end) {
    if (!(id %in% seen_ids)) {
      seen_ids <<- c(seen_ids,id)
      count <<- count + 1L
    }
    countall <<- countall+ 1L
    
    if ((snp_end & (uniqueN(seen_ids) >= thresh))|domain_end) {
      count <- count # copy enclosed value locally
      countall <- countall
      seen_ids <<- NULL
      count <<- 0L
      countall <<- 0L
    }
    paste(count,countall,sep=",")
  }
  unlist(Map(adjust_count, ID, domain_end, snp_end))
}


create_leaf_attribute <- function(snp_mat,snp_leaf_map){
  
  snp_mat@Dimnames <- list(NULL,NULL)
  
  #Convert mat to dgTMatrix if not already
  if(!inherits(snp_mat,"dgTMatrix")){
    snp_mat <- as(snp_mat, "dgTMatrix")
  }
  
  snp_mat2 <- snp_mat #copy object
  
  #Replace column indices with new set of indices
  #Make sure initial indices start with zero
  snp_mat2@j <- as.integer(snp_leaf_map[.(snp_mat@j+1)]$L1-1)
  #Correct dimensions of new matrix
  smij <- distinct_all(data.frame(snp_mat2@i,snp_mat2@j))
  snp_mat2@i <- smij[,1]
  snp_mat2@j <- smij[,2]
  snp_mat2@Dim <- as.integer(c(nrow(snp_mat2),
                               length(unique(snp_leaf_map$L1))))
  
  #Convert to dgCMatrix
  y <- as(snp_mat2,"dgCMatrix")
  return(y)
}

stouffer.p <- function(x){
  pnorm(sum(x)/sqrt(length(x)),lower.tail=FALSE)
}

stouffer.p2 <- function(y,g,wt){
  x <- qnorm(1-y)
  out=pnorm(sum(x*wt)/sqrt(uniqueN(g)),lower.tail=FALSE)
  return(rep(out,length(x)))
}




samp.pvals.leaf_L1 <- function(struct_map,Ll,Nsamp=100,p1s=NULL,p0s=NULL){
  # change to a more efficient algorithm
  # sampling (check the probability of the entile mixed population)
  delname <- intersect(c("snpID","alt_snps"),colnames(struct_map))
  struct_map2=struct_map%>%group_by_at(Ll)%>%
    mutate(mixed=((prod(hatm1)+prod(!hatm1))==0),
           altprop=sum(!hatm1)/length(hatm1),#altprop here means the estimated proportion of false rejection
           altct=sum(hatm1),
           nullct=sum(!hatm1))%>%ungroup()%>%
    dplyr::select(-delname)%>%distinct_all()
  if(any(sum(struct_map2$mixed)==0,length(p1s)==0)){return(NULL)}else{
    struct_map2 <- struct_map2%>%filter(mixed)%>%arrange(hatm1)
    struct_map3 <- struct_map2%>%group_by(altct,nullct,altprop)%>%
      summarize(ct=n())%>%ungroup()%>%rowid_to_column(var="type")
    sp1s <- sample(p1s,Nsamp*sum(struct_map3$altct),replace=TRUE)
    sp0s <- sample(p0s,Nsamp*sum(struct_map3$nullct),replace=TRUE)
    Sp1s <- data.frame(matrix(qnorm(sp1s,lower.tail=FALSE),ncol=Nsamp))%>%
      mutate(type=rep(struct_map3$type,time=struct_map3$altct))
    Sp0s <- data.frame(matrix(qnorm(sp0s,lower.tail=FALSE),ncol=Nsamp))%>%
      mutate(type=rep(struct_map3$type,time=struct_map3$nullct))
    sps <- rbind(Sp1s,Sp0s)%>%group_by(type)%>%
      summarise_at(paste0("X",seq(Nsamp)), stouffer.p)%>%ungroup()%>%
      left_join(struct_map3,by="type")
    return(list("sps"=sps,"M"=nrow(struct_map2),
                "altprop"=struct_map2$altprop))
  }
}
```

## The DYNATE Model

The `DYNATE` model and all the steps taken in the software are described in detail in our publication (Li, X., Pura, J., Allen, A., Owzar, K., Harms, M., & Xie, J. (2022)). It splits the whole-exome or genome into small regions called leaves, tests their disease associations, hierarchically aggregates these regions into larger regions, and then tests those larger regions. Here, we provide notations and include the formula and descriptions in this section as well.

After leaf P-values are calculated, we use Stouffer’s method to calculate node P-values. For node S, the P-value is
$$
T_s=\bar{\Phi}\left(\sum_{i=s} \bar{\Phi}^{-1}\left(T_i\right) / \sqrt{|S|}\right)
$$

where $\bar{\Phi}(·)$ is the complementary cumulative density function of the standard Gaussian distribution.

When a node is the union of multiple leaves, we might not be able to dichotomize its null and alternative status. For example, suppose node S = {1, 2} with leaf {1} is null and leaf {2} is alternative. Then $S$ is $1 / 2$ alternative. More generally, for any node $S$, let $S_1=S \cap\{i$ : $H_i$ is alternative $\}$. Then $\theta_S=\left|S_1\right| /|S|$. We refer to $S$ as $\theta_S$-alternative or $\left(1-\theta_S\right)$-null. To address this fractional null/alternative status of nodes, we introduce a new type I error measure called node-false-discovery-rate (node-FDR),
$$
\text { node-FDR }=\frac{\sum_{S \in \mathcal{R}}\left\{1-\theta_S\right\}}{|\mathcal{R}| \vee 1}
$$
Where $\mathcal{R}$ is the set of the rejected nodes. 

Firstly, we use `Test_leaf` function to test the leaf-level disease associations by generating leaf p-values for case-control study. Then use `DYNATE` function to conduct hierarchical multiple testing based on the leaf p-values. 

# Test_leaf Function

First, we introduce basic definition and set the tunning parameters as follows. 

Leaf: We define a leaf as a gentic region where $M$ subjects carry qualifying rare variants.

Leaf size: In a leaf, if there are $M$ subjects carry qualifying rare variants. Here, $M$ denotes the leaf size. It is a much smaller number compared to $N_0$ (Number of subjects assigned in case group) and $N_1$(Number of subjects assigned in control group). When $M = 1$, only one subject carries qualifying RVs in each leaf. In practice,we recommend $M \geq 5$. A node $S$ is a union of one or more leaves. A node represents a genome region with qualifying rare variants. On higher layers ($layer \geq 2$), it aggregates the accepting nodes from the previous layer to form new nodes. For any node $S$, the set of its qualifying RVs is denoted by $V(S)$. 

```{r parameters}
L <- 3 # layer number
M <- 5 # leaf size
alpha <- 0.05 # desired FDR
c=left_join(snp_info,sample_info,by="Sample.Name")
c$snpID=as.numeric(c$snpID)
c=c %>% arrange(snpID,Sample.Name,domainID)
```

Second, we use `Test_Leaf` function to construct leaves and generate leaf P-values for the case-control study.

About the parameter in function `Test_leaf`.

`snp_dat`: An data frame containing patients mutation information. The structure of it is showed above.

`thresh_val`:  A positive integer for leaf size.

`covars`: An optional vector about the name of covariates to be considered in the fitting process. Should be NULL (default) or a character vector.

`teststat`: The statistic used to derive p-value. Must be one of "FET" (Lancaster’s mid-P correction for the Fisher’s exact test) or "score" (efficient score statistics with saddle point approximation).

Construct leaves and generate leaf p-value based on score test statistics considering covariates effects.
```{r p_leaf.1}
p_leaf <- Test_Leaf1(snp_info = snp_info,sample_info = sample_info, thresh_val = M,teststat = "FET")
dim(p_leaf)
```
The dimension of the data is $66046\times 5$.

This output data frame has 66046 rows and 6 columns with name `L1`, `pvals`, `snpID`, `domainID`, `Test`. 
Each row links to a rare variant (SNPs) that belongs to a leaf with testing p-value < 1.

Here is the first 10 rows of the data.
```{r p_leaf.2}
head(p_leaf,10)
```

`L1`: Discrete numeric variable. Leaf ID. It indicates the leaf to which the SNP corresponding to this row is assigned. The `p_leaf` data contains `r length(unique(p_leaf$L1))` Leaf. 15826 < 66046 (the number of snp ID). This phenomenon illustrates that some leaves contain more than one rare variants (SNPs). The plot below gives us a more intuitive impression.

A histogram below plots the number of different rare variants(SNPs) each leaf has.
For example, when the x axis Leaf ID is 1, the y axis is 4, which means we observed 4 rare variants in leaf 1 genic region.

```{r domainID.2, echo=FALSE}
p4 = ggplot(data.frame(y = p_leaf$L1), aes(x = y)) +
  geom_histogram(binwidth = 0.2, color = "black") +
  labs(title = "The Number of Different Rare Variants(SNPs) Each Leaf Has", x = "Leaf ID", y = "Number of Snp ID") + 
  theme_bw()
p4
```

`pvals`: Continuous numeric variable. The leaf level p values. 

If it is less than threshold we calculated, then we call this leaf is null. The rare variants in this leaf are not associated with the disease; otherwise, we call it alternative.

The histogram below shows the distribution of p values. 

```{r pvals}
p5 = ggplot(data.frame(y = p_leaf$pvals), aes(x = y)) +
  geom_histogram(binwidth = 0.005) + scale_fill_manual(name = "p-values", 
                    values = "black") +
  labs(title = "Histogram of P-Values for All Leaves", x = "P-Values", y = "Frequency") + 
  theme_bw()
p5
```

`snpID`: Discrete numeric variable. In total there are `r length(unique(p_leaf$snpID))` SNPs. The meaning is the same as `raw`$snpID`above.

`domainID`: Discrete numeric variable. The generated data contains 2,000 domain IDs from 1 to 2000. The meaning is the same as `dat$domainID`above.

`Test`: It indicates the name of the statistical test to generate the leaf level p values (FET(Fisher’s exact test) or score(efficient score statistics with saddle point approximation)). In this study, we only use `FET` statistics.

# DYNATE Function
Finally, we use the function `DYNATE` to conduct dynamic and hierarchical testing based on the leaf level p values.

About the parameter in function `DYNATE`.

`struct_map`: A data frame with both leaf information and P-value information. We can see `p_leaf` above as an example.

`L`:	Maximum number of layers

`alpha`: Desired FDR (Significance level).

```{r dynate.1}
out <- DYNATE6(struct_map=p_leaf,L=L,alpha=alpha)
dim(out)
```
The dimension of the data is $9\times 10$.

This output data frame has 9 rows and 10 columns with name `L1`, `snpID`, `domainID`, `Test`, `pvals1`,`Layer`,`pvals-2`,`L2`,`pvals-3`,`L3`.

Each row links to a rare variant (SNPs) that belongs to a leaf with testing p-value < 1?

Here is the first 10 rows of the data.
```{r dynate.2}
tail(out,10)
p_leaf[p_leaf$L1 == 14656, ]
```

`pvals1`: Continuous numeric variable. Same as `pvals`. The node level p values.

`Layer`: Discrete numeric variable. It indicates in which layer the SNP is detected.
